---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# FindIT2

<!-- badges: start -->
<!-- badges: end -->

The goal of `FindIT2` is to ...

## Installation instructions

`FindIT2` is available on Bioconductor, you can install it by:
```{r 'install', eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install("FindIT2")
```

> For the packages on bioconductor, please make sure you download the latest stable `R` release from
[CRAN](http://cran.r-project.org/)

If you want the development version, install it directly from [GitHub](https://github.com/shangguandong1996/FindIT2):
```{r 'install_dev', eval = FALSE}
BiocManager::install("shangguandong1996/FindIT2")
```


## Document
If you want to download development version and view vignettes using `browseVignettes(FindIT2)`, your R version should be 4.0 or greater according to this [issue](https://github.com/Bioconductor/BiocStyle/issues/78) because I use the BiocStyle to output rmarkdown.
Then you have to firstly download the below packages
```{r, eval=FALSE}
packages <- c("BiocStyle", "knitr", "rmarkdown", "sessioninfo", "TxDb.Athaliana.BioMart.plantsmart28")
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
    BiocManager::install(packages[!installed_packages])
}
```
Meanwhile, adding the `build_vignettes=TRUE` when downloading development version
```{r, eval=FALSE}
BiocManager::install("shangguandong1996/FindIT2", build_vignettes=TRUE)
```
To view documentation of FindIT2, start R and enter:
```{r, eval = FALSE}
browseVignettes("FindIT2")
```
